---
title: "Integrating c4r with the R package toolchain"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Integrating c4r with the R package toolchain}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(c4r)
```

c4r is designed to slot into the existing R package-development toolchain
(devtools, usethis, roxygen2, knitr/rmarkdown, pkgdown, RStudio, GitHub
Actions). This vignette walks through each integration point.

## 1. Scaffolding with `use_c4r()`

In any R package directory, `use_c4r()` adds:

- `vignettes/architecture.Rmd` — a starter article
- an entry under `articles:` in `_pkgdown.yml`
- (optionally) a c4r chunk in `README.Rmd`
- a copy of c4r's pkgdown CSS

```r
use_c4r(vignette = TRUE, pkgdown = TRUE, readme = FALSE)
```

This mirrors the `usethis::use_*` family. Run it once; re-running is safe
because the helpers are idempotent (use `overwrite = TRUE` to replace).

## 2. A diagram from your package's `DESCRIPTION`

`c4_from_package()` reads `DESCRIPTION` and turns Imports/Depends into
external systems and your package into a central system. At the
container level, each `R/*.R` file becomes a container.

```{r, eval = FALSE}
c4_from_package(".") |> build_context()
c4_from_package(".", level = "container") |> build_container()
```

## 3. Declarative chunks via the `c4` knitr engine

c4r registers a knitr engine on load, so you can write diagrams in a
small DSL instead of R code:

````
```{c4, type="container", title="My App"}
person: user "End User"
container: api "API" "REST endpoint" [Service]
container: db "Database" "App data" [Postgres]
rel: user -> api "Uses" [HTTPS]
rel: api -> db "Reads/Writes" [SQL]
```
````

Renders as:

```{c4, type="container", title="My App"}
person: user "End User"
container: api "API" "REST endpoint" [Service]
container: db "Database" "App data" [Postgres]
rel: user -> api "Uses" [HTTPS]
rel: api -> db "Reads/Writes" [SQL]
```

The engine works in both R Markdown and Quarto.

## 4. pkgdown polish

Diagrams returned from `c4_*_diagram()` are htmlwidgets. They render
fine inline, but pkgdown's Bootstrap 5 layout works best when widgets
have an explicit width. `pkgdown_diagram()` sets sensible defaults:

```{r, eval = FALSE}
c4_from_template("three_tier") |> pkgdown_diagram(height = "600px")
```

`use_c4r_pkgdown()` also installs `pkgdown/extra.css` with responsive
SVG sizing.

## 5. RStudio addins

When c4r is installed, two entries appear in the RStudio addin menu:

- **Insert C4 chunk** — inserts a starter `c4` knitr chunk at the cursor.
- **Insert C4 template** — prompts for a template name and inserts an R
  chunk that renders it.

## 6. Continuous rendering with GitHub Actions

`use_c4r_action()` drops a reusable workflow into
`.github/workflows/c4r-render.yaml`. On every push it renders an SVG
of `c4_from_package()` and uploads it as a build artifact, so reviewers
can inspect architecture changes directly from a PR.
